---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ch.fso.besta.vacancies

The ch.fso.besta.vacancies package provides versioned time series data
and their meta information for scientific research.
In addition, the package contains the
extract-transform-load (ETL) functionality that
sources the data from its original provider.

## Browse Time Series Data

You can use GitHub's ability to render csv to explore the datasets.

## Installation

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("opentsi/opentimeseries")
```

## Basic Usage

Given a unique time series identifier and a GitHub repo,
*opentimeseries* will return a time series and long format `data.table`.

```{r example, eval=FALSE}
library(opentimeseries)
a <- read_open_ts("596.1",
                  remote_archive = "opentsi/ch.fso.besta.vacancies") # total job vacancies
a
```

By specifying a date in addition, you can able to obtain other versions
but the most recent one. The *opentimeseries* package will simply select
the most recent release that was available at the selected date.

```{r eval=FALSE}
a202307 <- read_open_ts("596.1",
                  remote_archive = "opentsi/ch.fso.besta.vacancies",
                  date = "2023-07-01")
```

Because time series data can get revised, storing vintages is important
to monitor data revisions and benchmark forecasts. Here's a quick visual
comparison:

```{r}
library(opentimeseries)
library(tsbox)

a <- read_open_ts("596.1",
                  remote_archive = "opentsi/ch.fso.besta.vacancies")
a202307 <- read_open_ts("596.1",
                  remote_archive = "opentsi/ch.fso.besta.vacancies",
                  date = "2023-07-01")

a202307$id <- sprintf("%s.202307", a202307$id)
a$id <- sprintf("%s.%s", a$id, Sys.Date())
ts_plot(rbind(a202307, a))
```
