---
title: "Automation of Dataset Updates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{automation_of_updates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Context:

To be able to automate the dataset updates, the data R package needs to be in place first. If that is not the case, follow the instructions in the vignette `initialise_a_ts_archive.Rmd`.

Now that you have a time series archive that contains an initial starting ground of versioned time series data, you need to care about getting regular updates of your data in automated fashion. The opentimeseries initiative (opentsi) designed a workflow to help you with that and to streamline the process. 

## Data Ingestion

First, a time series archive, that **was initiated with the help of deloRean package** provides you with a template to consume data from your data provider. This template leverages long time experience in data sourcing from public data providers. For example, it reminds you to implement a strategy to obtain and compare publication dates of from your data source. To find out more Tipps on this, go to the vignette: `write_your_process_data.Rmd`. 

## Scheduling Update Information

Second, your boilerplated data package, in our case `ch.kof.globalbaro` is equipped with a single entry point to use with your favorite scheduler:

```{r}
# your repo name
ch.kof.globalbaro::handle_update()
```

<!-- Assuming you have read the above vignette and implemented your own process_data.R function... -->

Running `handle_update` will trigger a comparison of the date of the last update of your time series version, and source data in case your version is outdated. The data ingestion process lives in the R folder of your data package, in our case `ch.kof.globalbaro/R`, as described in the vignette `write_your_process_data.Rmd`. 

## Github Actions Example 

Third, the boilerplated data package contains an example workflow for automatically running updates of your dataset using GitHub Actions (see `.github/workflows/update_data.yaml`). 
Of course you can use the entrypoint update function provided by your data package with any other CI/CD tools such as crow, woodpecker, Gitlab CI/CD or Azure DevOps pipelines. 
Let's break down the GitHub actions example workflow though in order to understand all steps: We use cron notation^[find more on cronjobs and cron notation at https://crontab.guru/] to let GitHub Actions know it should start the processing daily at 8.15 a.m and 11:15 a.m.

```yaml
name: Update Data
on:
  schedule:
    - cron:"15 8,11 * * *"
```

That is, spin up a VM and container, run the process in that container, push the artefact back up to Github, shut down the container and finalize the job. Note how GitHub runs docker inside a Virtual Machine and allows you to specify both the VM image and the container image. This a mere detail and mainly mentioned here to make sure the ubuntu image is not mistaking for container image here. Also, other container first CI/CD approaches do not set a VM images in their workflow files. 


```yaml
jobs:
  setup:
    runs-on: ubuntu-22.04-arm
    container: devxygmbh/r-alpine:4-3.21
    steps:
      - name: Manual Checkout
        run: |
          apk update
          apk add git
          git clone --depth=1 https://github.com/${{ github.repository
          }}.git .
          apk add harfbuzz-dev fribidi-dev

```

Note the use of ARM and alpine images here. Alpine is often called "the operating system of CI/CD" and is known for it's slim, resourcing saving image. 
Those more bare bone images require you to install system packages.
and to clone your repository to the container. 
We install the `pak` R package for package management and use it to install `opentimeseries` directly from its GitHub repository because our dataset package depends on opentimeseries which is not on CRAN yet. Hence `pak`  can not resolve the dependency automatically through the R ecosystem. 



```yaml

      - name: Configure Git
        run: |
          git config --global user.name "github-actions[bot]"
          git config --global user.email "github-actions[bot]@users.noreply.github.com"
          git config --global --add safe.directory $(pwd)


      - name: Set Repo
        run: |
          R -e 'options("repos")'
      - name: Install packages
        run: |
          R -q -e 'install.packages("pak")'
          R -q -e 'pak::pkg_install("remotes")'
          R -q -e 'remotes::install_github("opentsi/opentimeseries")'

```

Once we got opentimeseries we can run the `handle_update` function which is part of our boilerplate generated data package, in our case `ch.kof.globalbaro`.
Once the update is done and it's part of the cloned repository that lives inside the actions runner container, we can add and commit the changes and eventually push the data update back up to the GitHub repository. 


```yaml

      - name: handle_update
        run: |
          R -q -e '${{ github.repository }}::handle_update()'
      - name: Push changes
        run: |
          git add -A
          git commit -m "update data"
      - name: Push changes
        env:
          GITHUB_TOKEN: ${{ secrets.GITHUB_TOKEN }}
        run: git push https://x-access-token:${GITHUB_TOKEN}@github.com/${{ github.repository }}.git
```




<!-- - GHA Example 
- Single Entrypoint for (any) Schedulers  -->


<!-- To be able to automatically update datasets, deloRean builds a github actions configuration into your data package. To inspect this, navigate to `.github/workflows/ci.yaml`. More on this later. For this github action config to run properly, you need to configure a couple of things.  -->

## Runtime Environment Prerequisites

<!-- are we scaring people with this? -->
### Schedule
- Set a [schedule](https://docs.github.com/en/actions/reference/workflows-and-actions/events-that-trigger-workflows?search-overlay-input=cron#schedule) which specifies when your dataset should be checked for updates. We use POSIX cron syntax for this. Some examples are:

“At 00:00 on every day-of-week from Monday through Friday.”

```yaml
on:
  schedule:
    - cron: "0 0 * * 1-5"
```

"At 00:00 on the first day of the month"

```yaml
on:
  schedule:
    - cron: "0 0 1 * *"
```
[Here](https://crontab.guru/examples.html) are a couple more common use cases.


### ARM R Image

the github actions configuration uses a lightweight ARM Image to be able to run R 
<!-- TODO: Anything else to mention here? -->


## Build data R package

Your data R package currently looks like this:

```
├── .github
│   ├── workflows
│   │   ├── ci.yaml
├── data-raw
│   ├── series.csv
├── DESCRIPTION
├── NAMESPACE
├── R
│   ├── handle_update.R
│   ├── is_update_needed.R
│   ├── process_data.R
│   ├── dataset_write.R # previously (`dataset_update.R`)
├── README.md
```

### End-User Written Functions

In the R folder are boiler plated R functions. Most importantly, you have `handle_update` which calls the functions: `is_update_needed` which compares the dataset's LAST_UPDATE (aka the date written in this file, which is the last time the dataset was updated) with the data publisher's last update, and if `is_update_needed` returns True, then `process_data` ensures that the dataset is downloaded and is processed s.t. it has the same format as needed, and finally: `dataset_write` writes the data to disk (and then the gha pushes the final changes).

### Inside process_data.R

This function (?) processes data into a form that allows to map keys to folders. It also does a catalogue update (where it checks which keys are in this data package, and if these need to be updated). For example: in case we get a new dataset into this data package i.e. ch.kof.globalbaro -> coincident & leading, not just leading. 
<!-- TODO: Also dataseet specific metainformation -->

```r
opentimeseries::key_to_path()
```
<!--
One recommendation is to use the opentimeseries R package's `key_to_path` function, and `path_to_key` function. The former uses the key, such as `ch.kof.globalbaro` and nests it into the archive structure. I.e. 

TODO: where is this archive structure used? for users building their own data structure, yes, but many don't -> use opentimeseries to fetch data

-->

```
├── ch
│   ├── kof
│   │   ├── globalbaro
```

### Inside dataset write 

Used to the the `dataset_update` function. it just writes the new time series to the directory, then done.


### Understand & Adapt your GHA config 

<!-- TODO: in archive_init we also write a helper function: run_update ?? outdated -->

To be able to use your newly created data package well, you need to understand, how the update process works well. We already teasered above, that it uses the github action configuration, and that your data package needs "end-user written functions". But let's put it all together. 


**HOW UPDATING WORKS**: In your github actions configuration you are using a scheduler (described above) which specifies when the github actions script is run. This script, when run: takes your git repository (= your data R package) and installs the necessary dependencies and R packages needed to update your data package, and then runs the function `handle_update` which you have written to your data package. Effectively, `handle_update` checks whether an update is needed (`is_update_needed`) and if True, calls `process_data` which then downloads, and processes the data so it is in the unified tsl format. Finally, the `write_open_ts` function (from opentimeseries) writes the new dataset version to disk, and then the github action handles the rest (=commit, push)


<!-- TODO: wrote this before discussion on 23.1. Is this still necessary -->
**Step-by-Step Explanation:**
- provide explanation of boilerplated GHA config
- specify cron job (if deloRean default doesn't work for you)
- specify which R pkgs you need to install


<!-- ### TODO: timeseries specific metadata information (write boilerplated & docs for end-users) -->
