---
title: "automation_of_updates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{automation_of_updates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please first initialise a ts archive -> follow that vignette first.

## Runtime Environment Prerequisites

- Hints for Scheduling 
  - last couple of days/ last weekday of the month
  - twice a (week)day

- ARM R Image

## Build data R package

Your data R package currently looks like this:

```
├── data-raw
│   ├── series.csv
├── DESCRIPTION
├── NAMESPACE
├── R
│   ├── handle_update.R
│   ├── is_update_needed.R
│   ├── handle_downloads.R
│   ├── process_data.R
├── README.md
```
### End-User Written Functions

In the R folder are boiler plated R functions. Most importantly, you have `handle_update` which calls the functions: `is_update_needed` which compares the dataset's LAST_UPDATE with the publisher's last update, and if `is_update_needed` returns True, then `handle_downloads` runs, and downloads the new data. Lastly, `process_data` ensure that the dataset has the same format as needed. 

### Inside process_data.R

process data into a form that allows to map keys to folders. Maybe also catalogue update (where all keys are left) (in case we get a new dataset into this data package i.e. ch.kof.globalbaro -> coincident & leading, not just leading). Also dataseet specific metainformation 

```{r}
opentimeseries::key_to_path()
```

### Understand & Adapt your GHA config 

- provide explanation of boilerplated GHA config
- specify cron job (if deloRean default doesn't work for you)
- specify which R pkgs you need to install

### TODO: timeseries specific metadata information (write boilerplated & docs for end-users)