---
title: "General Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{General Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette describes the complete workflow for creating and publishing a time series archive using deloRean and the Open Time Series Initiative (opentsi).

## Workflow Overview

```
┌─────────────────────────────────────────────────────────────────────────────┐
│                         deloRean Workflow                                   │
└─────────────────────────────────────────────────────────────────────────────┘

     LOCAL DEVELOPMENT                          GITHUB / OPENTSI
     ─────────────────                          ─────────────────

    ┌─────────────────┐
    │  1. Init        │
    │     Archive     │
    └────────┬────────┘
             │
             ▼
    ┌─────────────────┐
    │  2. Import      │
    │     History     │
    └────────┬────────┘
             │
             ▼
    ┌─────────────────┐
    │  3. Add Update  │
    │     Process &   │
    │     Metadata    │
    └────────┬────────┘
             │
             │                              ┌─────────────────┐
             │  ─ ─ ─ ─ ─ ─ ─ ─ ─ ─ ─ ─ ─ ▶ │  4. Push to     │
             │       create repo first      │     GitHub      │
             │       on github.com/new      │  (your account) │
             │                              └────────┬────────┘
             │                                       │
             ▼                                       ▼
    ┌─────────────────┐                     ┌─────────────────┐
    │  Test & Refine  │◀ ─ ─ ─ ─ ─ ─ ─ ─ ─ ─│  GitHub Actions │
    │  Locally        │      CI feedback    │  runs updates   │
    └─────────────────┘                     └────────┬────────┘
                                                     │
                                                     ▼
                                            ┌─────────────────┐
                                            │  5. Request     │
                                            │     Listing     │
                                            └────────┬────────┘
                                                     │
                                                     ▼
                                            ┌─────────────────┐
                                            │  6. Transfer    │
                                            │     to opentsi  │
                                            └─────────────────┘

```

### Summary

| Step | Action | Documentation |
|------|--------|---------------|
| 1 | Initialize archive | `vignette("initialise_a_ts_archive")` |
| 2 | Import history | `vignette("initialise_a_ts_archive")` |
| 3 | Add update process | `vignette("write_your_process_data")`, `vignette("automation_of_updates")` |
| 3 | Add metadata | `vignette("meta_information")` |
| 4 | Push to GitHub | [github.com/new](https://github.com/new) |
| 5 | Request listing | [GitHub Issue](https://github.com/opentsi/opentsi/issues/new) |
| 6 | Transfer repository | [GitHub Docs: Transferring a repository](https://docs.github.com/en/repositories/creating-and-managing-repositories/transferring-a-repository) |


## Step 1: Initialize Archive

Create a new time series archive using `archive_init()`. This sets up the R package structure with all necessary boilerplate files.

```{r}
library(deloRean)

# Create a new archive
archive_init(
  archive_name = "ch.provider.dataset",
  parent_dir = "~/opentsi/"
)
```

This creates:

```
ch.provider.dataset/
├── .git/
├── .github/workflows/
│   └── update_data.yaml
├── data-raw/
│   └── metadata.yaml
├── inst/
├── R/
│   ├── handle_update.R
│   ├── is_update_needed.R
│   └── process_data.R
├── DESCRIPTION
├── NAMESPACE
└── README.md
```

See `vignette("initialise_a_ts_archive")` for details.


## Step 2: Import Historical Data

If you have historical vintages of your time series, import them with proper timestamps using `archive_import_history()`. Each vintage gets its own commit with the correct release date.

```{r}
# Prepare your vintages as a nested data.table
vintages_dt <- create_vintage_dt(release_dates, time_series_list)

# Import with historical commit dates
archive_import_history(
  vintages_dt,
  repository_path = "~/opentsi/ch.provider.dataset/"
)
```

This creates a git history where each commit represents a data release at its actual release date.


## Step 3: Add Update Process and Metadata

### Edit the Update Scripts

Customize the boilerplate R scripts in your archive:

- **`R/process_data.R`**: Fetch and transform data from the provider
- **`R/is_update_needed.R`**: Contains `compute_source_checksum()` - customize this to define how to detect source changes
- **`R/handle_update.R`**: Orchestrates the update process (usually no changes needed)

See `vignette("write_your_process_data")` and `vignette("automation_of_updates")` for details.

### Complete the Metadata

Edit `data-raw/metadata.yaml` with your dataset's information:

```yaml
country: "ch"
provider: "provider_name"
dataset: "dataset_name"
title:
  en: "Dataset Title in English"
  de: "Datensatz Titel auf Deutsch"
source_name:
  en: "Data Provider Name"
source_url: "https://provider.example.com/data"
dataset_frequency: "M"  # D, W, M, Q, Y, or I
```

Validate and render the metadata:

```{r}
validate_metadata("~/opentsi/ch.provider.dataset/data-raw/metadata.yaml")
render_metadata(repo_path = "~/opentsi/ch.provider.dataset/")
```

<!-- ### Initialize Checksum

After importing history and rendering metadata, establish the baseline checksum.
This prevents the first automated `handle_update()` from re-fetching data already in history:

```{r}
# Run from within your archive directory
init_checksum()
```
 -->

See `vignette("meta_information")` for the complete metadata specification.


## Step 4: Register Remote and Push to GitHub

### Create the GitHub Repository

1. Go to [github.com/new](https://github.com/new)
2. Enter your archive name as repository name (e.g., `ch.provider.dataset`)
3. Choose **Public** (recommended for open data) or Private
4. **Do NOT** initialize with README, .gitignore, or license (your archive already has these)
5. Click **Create repository**

### Push Your Archive

```{r}
archive_push_remote(
  repo_path = "~/opentsi/ch.provider.dataset",
  remote_owner = "your_github_username"
)
```

For SSH (default), ensure your SSH keys are configured with GitHub. For HTTPS:

```{r}
archive_push_remote(
  repo_path = "~/opentsi/ch.provider.dataset",
  remote_owner = "your_github_username",
  protocol = "https"
)
```

### Enable GitHub Actions

Go to your repository on GitHub → **Actions** tab → Enable workflows if prompted.

The included workflow will automatically run your update process on a schedule.


## Step 5: Request Listing in opentsi

Once your archive is working and regularly updating, request to be listed in the opentsi catalog:

1. Go to [github.com/opentsi/opentsi](https://github.com/opentsi/opentsi) (or the appropriate registry repo)
2. Open an issue using the "Dataset Listing Request" template
3. Provide:
   - Link to your repository
   - Brief description of the dataset
   - Confirmation that metadata is complete and valid
   - Your contact information

An opentsi maintainer will review your submission.


## Step 6: Transfer Repository and Become a Maintainer

After your listing request is approved:

### Transfer Your Repository to opentsi

1. Go to your repository → **Settings**
2. Scroll to **Danger Zone** → **Transfer repository**
3. Enter `opentsi` as the new owner
4. Confirm the transfer

An opentsi admin will accept the transfer. Your repository will move to `github.com/opentsi/ch.provider.dataset`, and the old URL will automatically redirect.

### Maintainer Access

After the transfer, opentsi admins will grant you maintainer access to your dataset repository so you can continue to:

- Update the data processing scripts
- Respond to issues
- Review contributions from others


## Workflow for opentsi Admins

If you are an opentsi admin creating an official dataset, you can skip the personal account step:

1. Create the repository directly under the opentsi organization
2. Push directly with `remote_owner = "opentsi"`:

```{r}
archive_push_remote(
  repo_path = "~/opentsi/ch.provider.dataset",
  remote_owner = "opentsi"
)
```


