---
# i propose an alternative title: host a ts-archive with existing vintages
# and the second vignette: host a ts-archive (getting started)
title: "hosting-an-own-ts-archive"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hosting-an-own-ts-archive}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Load the necessary packages for this reprex


```{r setup}
library(kofdata)
library(data.table)
library(tsbox)
library(devtools)

# TODO: can't install because no namespace export failed
remotes::install_github("opentsi/deloRean")
library(deloRean)
```


To initialise the repo, input name of new repo, and location to put it, and set env. vars.


```{r initialise repo}
archive_init("ch.kof.globalbaro", "~/KOF_Lab/opentsi/")

Sys.setenv("DELOREAN_TZ" = "UTC")
Sys.setenv("DELOREAN_EMAIL" = "bannert@kof.ethz.ch")
Sys.setenv("DELOREAN_AUTHOR" = "Open Time Series Initiative")
Sys.setenv("DELOREAN_SSH_KEY" = "~/.ssh/id_rsa")
```

This creates a new R package in the specified location. Check it out by navigating to that location, and inspecting the files. You should see the following directory structure under `ch.kof.globalbaro`:

```
├── data-raw
├── DESCRIPTION
├── NAMESPACE
├── R
├── README.md
```

After this, we can initialise our mock vintages dataframe.


```{r}
global <- kofdata::get_collection("globalbaro_vintages")
names(global) <- gsub("globalbaro_","",names(global))
names(global) <- sub("_", "\\.", names(global))
class(global) <- c(class(global), "tslist")

release_dates <- rep(seq(as.Date("2020-01-10"),
                     by = "1 month",
                     length.out = length(global)/2),2)

vintages_dt <- create_vintage_dt(release_dates, global)
head(vintages_dt)
```

The Result is a nested data frame, with the variables: `id`, `release_date` and `data`. In `data` you will find the time series data, with the values `time` and `value`.

Now, all we have to do is run the `deloRean::archive_import_history` function, which creates a git history for the entire vintages, and assigns each vintage its respective release date

```{r backlog}
archive_import_history(vintages_dt, "../ch.kof.globalbaro/")
```

To check out, what this does, navigate to your data package, in our case `ch.kof.globalbaro` and run `git log`. Then you will see the entire vintages and their commit history. 

Once you have initialised the data package and added the commit history, you can put inplace the processing script, which is responsible for re-fetching the new data, this should be written in the `process_data.R` file. This file could look as follows:

```{r}
# TODO: write `process_data.R` file for our ch.kof.globalbaro example
# check LAST_UPDATE 
# fetch data from data provider (via API e.g.)
# clean/process data 
```

Then, the `process_data.R` file is used for the `dataset_update` function (which is automatically run, every xyz days, e.g.) 
