---
title: "Initialising a Time Series Archive"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{initialise_a_ts_archive"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- TODO: check if this does global eval = FALSE -->
```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Load the necessary packages for this reprex


```{r setup}
library(remotes)

install_github("opentsi/opentimeseries",ref = "experiment/opencode")
install_github("opentsi/deloRean", ref = "cleanup")
library(deloRean)
```


To initialise the repo, input name of new repo, and location to put it, and set env. vars.


```{r initialise repo}
archive_init("ch.kof.globalbaro", parent_dir = "~/KOF_Lab/opentsi/")

Sys.setenv("DELOREAN_TZ" = "UTC")
Sys.setenv("DELOREAN_EMAIL" = "bannert@kof.ethz.ch")
Sys.setenv("DELOREAN_AUTHOR" = "Open Time Series Initiative")
Sys.setenv("DELOREAN_SSH_KEY" = "~/.ssh/id_rsa")
```

This creates a new R package in the specified location. Check it out by navigating to that location, and inspecting the files. You should see the following directory structure under `ch.kof.globalbaro`:

```
├── data-raw
├── DESCRIPTION
├── NAMESPACE
├── R
│   ├── handle_update.R
│   ├── is_update_needed.R
│   ├── process_data.R
├── README.md
```

After this, we can initialise our mock vintages dataframe.

Let's load a couple of packages to get and label our demo dataset.

```{r pre-process}
library(kofdata)
library(data.table)
library(tsbox)

global <- kofdata::get_collection("globalbaro_vintages")
names(global) <- gsub("globalbaro_","",names(global))
names(global) <- sub("_", "\\.", names(global))
class(global) <- c(class(global), "tslist")

release_dates <- rep(seq(as.Date("2020-01-10"),
                     by = "1 month",
                     length.out = length(global)/2),2)

```

Now that we have a nested list of vintages, and corresponding releasedates, we can use deloRean's `create_vintage_dt` function to create a nested data.table containing multiple versions of multiple time series (here coincident, leading)

```{r}
vintages_dt <- create_vintage_dt(release_dates, global)
head(vintages_dt)
```

Our nested data frame has the variables: `id`, `release_date` and `data`. In `data` you will find the time series data, in the columns `time` and `value`.

Now, all we have to do is run the `deloRean::archive_import_history` function, which creates a git history for the entire set of vintages, and assigns each vintage its respective release date.

```{r backlog}
archive_import_history(vintages_dt, repository_path = "../ch.kof.globalbaro/")
```

To check out what this does, navigate to your data package, in our case `ch.kof.globalbaro` and run `git log`. Then you will see the entire vintages and their commit history. 

Your git history should look something like this:

```{sh}
commit 3185998655bc219d8c907d94dd3baf7c78a4a703
Author: Open Time Series Initiative <bannert@kof.ethz.ch>
Date:   Thu Dec 11 00:59:59 2025 +0100

    opentsi full history init

commit 2e38880f59b7f802595fe2b6a200263d6c3de106
Author: Open Time Series Initiative <bannert@kof.ethz.ch>
Date:   Tue Nov 11 00:59:59 2025 +0100

    opentsi full history init

commit 5dac427822f081d9986152f1d7bbb13214877266
Author: Open Time Series Initiative <bannert@kof.ethz.ch>
Date:   Sat Oct 11 01:59:59 2025 +0200

    opentsi full history init

commit 09ef30ca21c854bc2770cd81e273286f61e43f98
Author: Open Time Series Initiative <bannert@kof.ethz.ch>
Date:   Thu Sep 11 01:59:59 2025 +0200

    opentsi full history init

commit e7721f4e9a50ebc2f99290c73a03c5e6e2032126
Author: Open Time Series Initiative <bannert@kof.ethz.ch>
Date:   Mon Aug 11 01:59:59 2025 +0200

    opentsi full history init

commit 50957e2160fb7b1fefd23b28e124102e91c68e33
Author: Open Time Series Initiative <bannert@kof.ethz.ch>
Date:   Fri Jul 11 01:59:59 2025 +0200

    opentsi full history init

commit 04acff142152d794d37fc05c77d77b5a6b35abd2
Author: Open Time Series Initiative <bannert@kof.ethz.ch>
```

<!-- TODO: include code block to see git log-->

<!-- 
Once you have initialised the data package and added the commit history, you can put inplace the processing script, which is responsible for re-fetching the new data, this should be written in the `process_data.R` file. This file could look as follows:

```{r}
# TODO: write `process_data.R` file for our ch.kof.globalbaro example
# check LAST_UPDATE 
# fetch data from data provider (via API e.g.)
# clean/process data 
```

Then, the `process_data.R` file is used for the `dataset_update` function (which is automatically run, every xyz days, e.g.)  -->


<!-- TODO: How does the repo look after this process -->