---
title: "Metainformation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Metainformation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette explains how to add metadata to your time series archive. Good metadata makes your data discoverable, understandable, and usable across languages and tools.

## Quick Start

After [initializing your archive](initialise_a_ts_archive.html), you'll find a metadata template at `data-raw/metadata.yaml`. The workflow is:

```{r}
library(deloRean)

# 1. Initialize archive (creates metadata template)
archive_init("ch.kof.globalbaro", parent_dir = "~/repositories/opentsi/")

# 2. Edit data-raw/metadata.yaml (fill in your dataset info)

# 3. Validate and render to JSON
render_metadata("~/repositories/opentsi/ch.kof.globalbaro")
```
```
#> Metadata validation passed.
#> Rendered: data-raw/metadata.yaml -> inst/metadata.json
```

The YAML file is your source (human-readable, editable). The JSON file is the artifact consumed by `opentimeseries`:

```{r}
library(opentimeseries)

# Read metadata (local or remote)
meta <- read_meta("~/repositories/opentsi/ch.kof.globalbaro")

# Get all keys
get_keys(meta)
#> [1] "ch.kof.globalbaro.leading"
#> [2] "ch.kof.globalbaro.coincident"

# Resolve to human-readable labels
resolve_meta(meta, lang = "en")
```

---

## Key Naming Convention

OpenTSI uses **speaking keys** - identifiers that are unique, memorable, and hierarchical:

```
country.provider.dataset.dimension.variable.unit
```

### Example

The KOF Swiss Economic Institute uses this pattern:

```
ch.kof.inu.ng08.fx.q_ql_ass_bs.balance
```

| Segment | Meaning |
|---------|---------|
| ch | Switzerland (ISO country code) |
| kof | KOF Swiss Economic Institute |
| inu | Manufacturing Survey |
| ng08 | NACE Sector 08 |
| fx | Mixed frequency |
| q_ql_ass_bs | Qualitative question, business assessment |
| balance | Balance (unit) |

**Benefits:**

- Column names are self-documenting during analysis
- Combinatorics: translate chunks once, generate labels for thousands of series
- Hierarchical: easy filtering (`grep("^ch.kof.inu", keys)`)

---

## Required Dataset Fields

Every metadata file must include these fields:

```yaml
# Identity (forms the key prefix: ch.kof.globalbaro)
country: "ch"           # ISO 3166-1 alpha-2 (lowercase)
provider: "kof"         # Short identifier (lowercase, a-z, 0-9, _)
dataset: "globalbaro"   # Short identifier (lowercase, a-z, 0-9, _)

# Descriptions (multilingual, English required)
title:
  en: "Global Economic Barometer"
  de: "Globales Wirtschaftsbarometer"

source_name:
  en: "KOF Swiss Economic Institute"
  de: "KOF Konjunkturforschungsstelle"

source_url: "https://kof.ethz.ch"

# Frequency: D=daily, W=weekly, M=monthly, Q=quarterly, Y=yearly, I=irregular
dataset_frequency: "M"

# Units (use "all" if uniform across variables)
units:
  all:
    en: "index"
    de: "Index"
```

**Note:** All series in a dataset must share the same frequency. Different frequencies = different datasets (repos).

---

## Flat vs Nested Datasets

### Flat Dataset (No Grouping)

For simple datasets with a few distinct series:

```yaml
dim_order: []  # empty - no grouping dimensions

labels:
  dimnames:
    leading:
      en: "Leading Indicator"
      de: "Frühindikator"
    coincident:
      en: "Coincident Indicator"
      de: "Gleichlaufender Indikator"
```

This produces keys:

```
ch.kof.globalbaro.leading
ch.kof.globalbaro.coincident
```

### Grouped Dataset (with dim_order)

When variables are grouped by dimensions (like sex, region):

```yaml
dim_order: ["region"]  # region is a grouping dimension

labels:
  dimnames:
    region:
      en: "Region"
    eu:
      en: "Europe"
    us:
      en: "United States"
    leading:
      en: "Leading Indicator"
```

The `dim_order` tells the system that `region` is a structural dimension, not a series name.

### Hierarchical Dataset (with hierarchy)

For datasets with aggregation structure (e.g., states to country totals):

```yaml
dim_order: ["residence", "sex"]

hierarchy:
  residence:
    tot:            # total
      ch: {}        # Swiss
      frgn:         # foreign
        res: {}     # residents
        seas: {}    # seasonal workers
  sex:
    tot:
      m: {}         # male
      f: {}         # female

aggregate:
  all: "sum"  # how to aggregate: mean, sum, first, last
```

This defines which series can be summed to produce totals.

---

## Labels

Labels provide human-readable names for key segments. Structure:

```yaml
labels:
  dimnames:
    # Dimension names (structural)
    idx_type:
      en: "Index Type"
      de: "Indextyp"

    # Series names (actual variables)
    leading:
      en: "Leading Indicator"
      de: "Frühindikator"
    coincident:
      en: "Coincident Indicator"
      de: "Gleichlaufender Indikator"
```

**How it works:**

- Everything under `dimnames` gets a label
- `dim_order` items are excluded from key generation (they're grouping dimensions)
- Remaining items become the series keys

**Multilingual support:**

- English (`en`) is required as the baseline
- Add any ISO 639-1 language codes as needed
- `resolve_meta(meta, lang = "de")` retrieves German labels, falls back to English if missing

---

## Validation and Rendering

### Step 1: Validate

Before rendering, validate your metadata:

```{r}
# Validates against the OpenTSI schema
validate_metadata("~/repositories/opentsi/ch.kof.globalbaro")
```

Common validation errors:

- Missing required fields
- Invalid country code (must be 2 lowercase letters)
- Invalid language codes (must be ISO 639-1)
- Invalid frequency (must be D, W, M, Q, Y, or I)

### Step 2: Render

Convert YAML to JSON:

```{r}
render_metadata("~/repositories/opentsi/ch.kof.globalbaro")
```

This:

1. Reads `data-raw/metadata.yaml`
2. Validates the content
3. Writes `inst/metadata.json`

---

## Complete Example

Here's a full `metadata.yaml` for the KOF Global Barometer:

```yaml
# === Dataset Identity ===
country: "ch"
provider: "kof"
dataset: "globalbaro"

# === Descriptions ===
title:
  en: "Global Economic Barometer"
  de: "Globales Wirtschaftsbarometer"

source_name:
  en: "KOF Swiss Economic Institute"
  de: "KOF Konjunkturforschungsstelle"

source_url: "https://kof.ethz.ch/en/forecasts-and-indicators/indicators/kof-globalbarometer.html"

# === Data Properties ===
dataset_frequency: "M"

units:
  all:
    en: "index"
    de: "Index"

aggregate:
  all: "mean"

# === Structure ===
dim_order: []

hierarchy: {}

# === Labels ===
labels:
  dimnames:
    leading:
      en: "Leading Indicator"
      de: "Frühindikator"
    coincident:
      en: "Coincident Indicator"
      de: "Gleichlaufender Indikator"
```

After rendering, consumers can:

```{r}
library(opentimeseries)

meta <- read_meta("opentsi/ch.kof.globalbaro")  # from GitHub

get_keys(meta)
#> [1] "ch.kof.globalbaro.leading"
#> [2] "ch.kof.globalbaro.coincident"

resolve_meta(meta, lang = "de")
#>                              key                                      label_full                    label_short
#> 1 ch.kof.globalbaro.leading      KOF Konjunkturforschungsstelle, Globales...    Globales Wirtschaftsbarometer, M, Frühindikator
#> 2 ch.kof.globalbaro.coincident   KOF Konjunkturforschungsstelle, Globales...    Globales Wirtschaftsbarometer, M, Gleichlaufender Indikator

# Print formatted
print_resolved(resolve_meta(meta), format = "short")
#> ch.kof.globalbaro.leading
#>   -> Global Economic Barometer, M, Leading Indicator
#>
#> ch.kof.globalbaro.coincident
#>   -> Global Economic Barometer, M, Coincident Indicator
```

---

## JSON Schema

The official schema is available at:

- Package: `inst/schema/opentsi_metadata.schema.json`
- Online: `https://opentsi.github.io/schema/v1/metadata.json`

View in R:

```{r}
deloRean::view_schema()
```

---
## Summary

| File | Location | Purpose |
|------|----------|---------|
| metadata.yaml | data-raw/ | Source (human-editable) |
| metadata.json | inst/ | Artifact (consumed by opentimeseries) |

| Function | Package | Purpose |
|----------|---------|---------|
| `validate_metadata()` | deloRean | Check YAML against schema |
| `render_metadata()` | deloRean | YAML to JSON |
| `read_meta()` | opentimeseries | Load metadata (local or remote) |
| `get_keys()` | opentimeseries | Extract all keys |
| `resolve_meta()` | opentimeseries | Map keys to labels |
