---
title: "Write your Process Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{write_your_process_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Context:

Assume you are a research assistant, tasked with downloading and processing the **latest** time series vintage, so that it can be later used for analyses. Assuming you already know the dataset you need to fetch, and the data source which publishes it. 

Let's go through this process, by assuming you want to fetch and process the [KOF's Global Barometer](https://kof.ethz.ch/en/forecasts-and-indicators/indicators/kof-globalbaro.html) which are a system of indicators that enables a timely analysis of global economic development. You know that the KOF Institute publishes this dataset, and you know that there are 2 main ways to fetch it. Let's go through them.

Assuming that you know that the global barometer is timeseries list, consisting of the leading and coincident globalbaro:


### Process Globalbaro Data - with {kofdata}

Since there exists a KOF API wrapper in R, you can use this to import and then process your data

<!-- check boilerplating from process_data in initialised folder. include opentimeseries key_to_path -->
```{r message= FALSE}
library(kofdata)

#' Process KOF Vintage Data into Time Series List (tsl)
#'
#' Fetches vintage data from the kofdata R package, filters for the most recent vintage of specific 
#' indicators, and cleans the resulting names.
#'
#' @param key A character string of the KOF set key to fetch.
#' @param ids A character vector of indicator IDs to extract (e.g., c("coincident", "leading")).
#'
#' @return A named list of ts objects (tsl) with cleaned names in the key.id format
#' @export
#'
process_data <- function(key, ids = NULL){
  # fetch initial tsl
  tsl <- get_collection(key)
  names <- names(tsl)

  # helper function used to get last vintage for each id
  last_vintages <- sapply(ids, function(id) {
  # use pattern to match the specific ID (e.g., "_coincident_")
  pattern <- paste0("_", id, "_")
  
  # filter all names of ts that match this
  matches <- names[grepl(pattern, names)]
  
  # Sort them (alphabetical works for YYYY-MM) and take the last one
  return(tail(sort(matches), 1))
  })
  res_tsl <- tsl[last_vintages]

  # since we return a tsl, all we need to do to finish processing is remove the suffix for each key 
  # (to write the data properly later)
  names(res_tsl) <- sapply(strsplit(names(res_tsl), "_"), `[`, 2)

  return(res_tsl)
}

tsl <- process_data("globalbaro_vintages", ids = c("coincident", "leading"))

# plot all ids together in one plot, so need to combine to matrix
ts_matrix <- do.call(cbind, tsl)
ts.plot(ts_matrix, col = 1:ncol(ts_matrix)) 
```


### Process Globalbaro Data - with the KOF API

Now assuming that we don't know about kofdata, then we can use the data providers API (if it is available). First, we read the [API's documentation](https://datenservice.kof.ethz.ch/) and then we use the httr2 package.

```{r}
library(httr2)

#' Fetch Raw JSON Data from KOF API
#'
#' Internal helper to execute GET requests against the KOF public sets API.
#'
#' @param key A character string representing the set key (e.g., "globalbaro_vintages").
#'
#' @return A list containing the parsed JSON response.
fetch_data <- function(key){
  # try pasting the base_set_url into your browser then you see all available sets
  set_url <- "https://datenservice.kof.ethz.ch/api/v1/public/sets/"
  get_url <- paste0(set_url, key)
  # print(get_url)

  response <- request(get_url) |> req_perform() 

  return(resp_body_json(response))
}

```


```{r}
#' Process KOF Vintage Data into Time Series List (tsl)
#'
#' Fetches vintage data from the KOF API, converts the JSON structure into 
#' ts objects, filters for the most recent vintage of specific 
#' indicators, and cleans the resulting names.
#'
#' @param key A character string of the KOF set key to fetch.
#' @param ids A character vector of indicator IDs to extract (e.g., c("coincident", "leading")).
#'
#' @return A named list of ts objects (tsl) with cleaned names in the key.id format
#' @export
#'
process_data <- function(key, ids = NULL){

  # fetch data 
  raw_json <- fetch_data(key)

  # turn json into tsl
  tsl <- lapply(raw_json, function(vintage) {
    
    # unlist and get all values, convert to numeric
    values <- as.numeric(unlist(lapply(vintage, function(obs) obs$value)))
    
    # Extract first observation, turn into date
    start_str <- vintage[[1]]$date
    start_date <- as.numeric(unlist(strsplit(start_str, "-")))
    
    # Create the ts object
    return(ts(values, start = start_date, frequency = 12))

  })

  # select last vintage for each timeseries id
  all_names <- names(tsl)

  last_names <- sapply(ids, function(id) {
    matches <- all_names[grepl(paste0("_", id, "_"), all_names)]
    tail(sort(matches), 1)
  })

  tsl_subset <- tsl[last_names]

  # clean names, to write_open_ts seamlessly later
  names(res_tsl) <- sapply(strsplit(names(res_tsl), "_"), `[`, 2)

  return(tsl_subset)

}

tsl <- process_data("globalbaro_vintages", ids = c("coincident", "leading"))

# plot all ids together in one plot, so need to combine to matrix
ts_matrix <- do.call(cbind, tsl)
ts.plot(ts_matrix, col = 1:ncol(ts_matrix)) 
```

This is how you could write a `process_data` function, which gets the globalbaro vintages, and also transforms them into the timeseries list format. 

Please remember to put this function into the process_data.R file, which was created upon initialising this archive. Follow the vignette `initialise_a_ts_archive.Rmd` if that is not the case. 

<!-- TODO: finish this example. still need to include opentimeseries::key_to_path() and mapping latest imported ts to said key_path -->


### Finishing Touches

But we are not quite done yet. Now you have seen how to get the latest time series vintages of ch.kof.globalbaro, but the process data step is only done when the data is written to its designated spot. In our case:

```
├── data-raw
│   ├── coincident
│   │   ├── series.csv
│   ├── leading
│   │   ├── series.csv
│   ├── index.csv
├── DESCRIPTION
├── NAMESPACE
├── R
│   ├── handle_update.R
│   ├── is_update_needed.R
│   ├── process_data.R
├── README.md
```

Our data needs to go into the respective `series.csv` files, of each sepeater sub-key.

To do this, we need to add the following code to our above implementation, before the final `return(res_tbl)` line:

<!-- TODO: where key_to_path usage??? -->

```{r}

# Write each time series to series.csv in the corresponding directory
lapply(names(res_tsl), function(key) {
  # get ts$key
  ts_obj <- res_tsl[[key]]

  # Convert ts object to data frame with time and value columns
  # Extract time index and values from the ts object
  values <- as.numeric(ts_obj)

  ts_dates <- as.Date(lubridate::date_decimal(as.numeric(time(ts_obj))))

  # Create data frame in the format of the input file
  ts_df <- data.frame(
    time = as.Date(ts_dates),
    value = values
  )

  # Create path to write file
  output_path <- file.path(getwd(), "data-raw", key, "series.csv")

  # Write to CSV without row names
  write.csv(ts_df, file = output_path, row.names = FALSE)
})

```

This code snippet ensures that for each key in the dataset (here, coincident, leading), the data for this sub-key is written in the correct format ,as previously written, to `series.csv`.