---
title: "write_your_process_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{write_your_process_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Context:
<!-- TODO: no reference to gha, no deloRean, etc. just teach people to process data, with roxygen -->

Assume you are a research assistant, tasked with downloading and processing the latest time series vintage, so that it can be later used for analyses. Assuming you already know the dataset you need to fetch, and the data source which publishes it. 

Let's go through this process, by assuming you want to fetch and process the [KOF's Global Barometer](https://kof.ethz.ch/en/forecasts-and-indicators/indicators/kof-globalbaro.html) which are a system of indicators that enables a timely analysis of global economic development. You know that the KOF Institute publishes this dataset, and you know that there are 2 main ways to fetch it. Let's go through them.

Assuming that you know that the global barometer is timeseries list, consisting of the leading and coincident globalbaro:


### Process Globalbaro Data - with {kofdata}

Since there exists a KOF API wrapper in R, you can use this to import and then process your data


<!-- TODO: include roxygen comments -->
```{r message= FALSE}
library(kofdata)

process_data <- function(key, ids = NULL){
  # fetch initial tsl
  tsl <- get_collection(key)
  names <- names(tsl)

  # helper function used to get last vintage for each id
  last_vintages <- sapply(ids, function(id) {
  # use pattern to match the specific ID (e.g., "_coincident_")
  pattern <- paste0("_", id, "_")
  
  # filter all names of ts that match this
  matches <- names[grepl(pattern, names)]
  
  # Sort them (alphabetical works for YYYY-MM) and take the last one
  return(tail(sort(matches), 1))
  })
  res_tsl <- tsl[last_vintages]

  # since we return a tsl, all we need to do to finish processing is remove the suffix for each key 
  # (to write the data properly later)
  names(res_tsl) <- sub("_", ".", sub("_[0-9]{4}-[0-9]{2}$", "", names(res_tsl)))

  return(res_tsl)
}

tsl <- process_data("globalbaro_vintages", ids = c("coincident", "leading"))

# plot all ids together in one plot, so need to combine to matrix
ts_matrix <- do.call(cbind, tsl)
ts.plot(ts_matrix, col = 1:ncol(ts_matrix)) 
```


### Process Globalbaro Data - with the KOF API

Now assuming that we don't know about kofdata, then we can use the data providers API (if it is available). First, we read the [API's documentation](https://datenservice.kof.ethz.ch/) and then we use the httr2 package.

```{r}
library(httr2)

# define a helper function to execute the get request and response
fetch_data <- function(key){
  # try pasting the base_set_url into your browser then you see all available sets
  set_url <- "https://datenservice.kof.ethz.ch/api/v1/public/sets/"
  get_url <- paste0(set_url, key)
  # print(get_url)

  response <- request(get_url) |> req_perform() 

  return(resp_body_json(response))
}

```


```{r}
process_data <- function(key, ids = NULL){

  # fetch data 
  raw_json <- fetch_data(key)

  # turn json into tsl
  tsl <- lapply(raw_json, function(vintage) {
    
    # unlist and get all values, convert to numeric
    values <- as.numeric(unlist(lapply(vintage, function(obs) obs$value)))
    
    # Extract first observation, turn into date
    start_str <- vintage[[1]]$date
    start_date <- as.numeric(unlist(strsplit(start_str, "-")))
    
    # Create the ts object
    return(ts(values, start = start_date, frequency = 12))

  })

  # select last vintage for each timeseries id
  all_names <- names(tsl)

  last_names <- sapply(ids, function(id) {
    matches <- all_names[grepl(paste0("_", id, "_"), all_names)]
    tail(sort(matches), 1)
  })

  tsl_subset <- tsl[last_names]

  # clean names, to write_open_ts seamlessly later
  names(tsl_subset) <- sub("_", ".", sub("_[0-9]{4}-[0-9]{2}$", "", names(tsl_subset)))

  return(tsl_subset)

}

tsl <- process_data("globalbaro_vintages", ids = c("coincident", "leading"))

# plot all ids together in one plot, so need to combine to matrix
ts_matrix <- do.call(cbind, tsl)
ts.plot(ts_matrix, col = 1:ncol(ts_matrix)) 
```

This is how you could write a `process_data` function, which gets the globalbaro vintages, and also transforms them into the timeseries list format. 

